Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation

### 2022-02-06 Writing/JournalSubmission/ML_VRA_format.rmd
Correct the structure of CF_for_VRA folder

```{r echo = F, results='hide', message=FALSE}
library(knitr)
# library(here)

# opts_knit$set(root.dir = here())
opts_knit$set(root.dir = "~/Dropbox/ResearchProject/ML_VRA")

@@

```{r, cache = FALSE}
# === packages ===#
library(sf)
library(png)
library(data.table)
library(grf)
library(RColorBrewer)
library(patchwork)
library(plotly)
library(magrittr)
library(ggplot2)
library(viridis)
library(tmap)
library(tidyverse)
library(ggthemes)
library(mgcv)
library(ggpubr)
library(flextable)
library(officer)
library(officedown)
library(modelsummary)
library(raster)
library(bookdown)
library(png)
library(grid)
```

```{r analyses, results = "hide", fig.show='hide', cache = FALSE}
#/*=================================================*/
#' # Preparation
#/*=================================================*/
# This is where you prepare data sets and results 
# library(here)
# setwd(here())
#/*----------------------------------*/
#' ## Create figures and tablesb
#/*----------------------------------*/
# --- Convert an Rmd file into a pure R file using purl() ---#
knitr::purl("./GitControlled/Codes/PrepareResults.Rmd", output = "./GitControlled/Codes/PrepareResults.R")

# --- Read the simulation results and prepare figures and tables ---#
source("./GitControlled/Codes/PrepareResults.R")
```

@@

@@

CF was developed specifically for identifying heterogeneous treatment effects [@athey2016recursive; @Wager2018a]. Like RF, CF works by building an ensemble of trees, called *causal trees* (CT), by recursively partitioning covariate space into two leaves, where each leaf estimates conditional average treatment effects as the mean differences between the dependent variable (here, yield) of the treated and control observations that belong to the leaf. Figure \@ref(fig:ct-ex) shows an example of CT where the treatment is an increase in seed rate from $27,000$ seeds/ha (control) to $31,000$ seeds/ha (treated) with apparent electrical conductivity ($\textit{ecs}$)($dS/m$) and topographical $\textit{slope}$ ($degree$) as variables to explain the heterogeneity of the impact of the treatment impact. 
In the first node, all the samples ($1024$ observations) belong to the node, and the causal effect of the treatment was calculated as the difference in the mean yield of the treated and control groups, which turned out to be $593$ kg/ha in this example.

```{r ct-ex, fig.width=4, fig.height=4, fig.align="center", fig.cap = "An illustrative example of causal tree"}
# knitr::include_graphics("./GitControlled/Writing/Re_JournalSubmission/ctree_ex.pdf")
readPNG("./GitControlled/Writing/Re_JournalSubmission/ctree_ex.png") %>% grid.raster()
```

`r ftext('NOTE: The first number in each node indicates the predicted treatment effects (in kg/ha). The number and the percentage of observations belonging to the node are shown below that.', fp_text(font.size = 9))`

@@

Monte Carlo simulations were conducted for nitrogen experiments for corn production to estimate the economically optimal nitrogen rate (EONR). The first step in our simulation was to generate one thousand synthetic fields that "resembled" an actual field. <span style="color:red"> After generating one thousand synthetic fields, we repeated to evaluate each of the ML methods by one thousand times. In an individual simulation round, one of the fields was used as a training data set. As a validation data set, a different field from a training data set in that simulation round was used.</span> <!-- For each simulated field, we generated a training data set and a validation data set. --> Field characteristics and yield were simulated based on the Mitscherlich-Baule (MB) yield response function. We used the RF, BRF, and CNN ML methods to estimate yield response functions, and we used the CF method to estimate treatment effects. For each method, we estimated site-specific EONRs using the training data set, and checked the statistical performance of the predicted site-specific EONR using the validation data set. Detailed descriptions of these steps are provided below. The R codes that implement the simulation analyses are available at https://github.com/tmieno2/ML_VRA.git.

@@

### Simulated Experimental Fields
<!-- experimental field: area: 51ha, plot: 384, subplot:1536, cell:55296   -->

The simulated field was created based on an actual 51-ha Illinois field on which the Data-Intensive Farm Management (DIFM) project [@bullock2019] has run multiple OFPEs. Figure \@ref(fig:field-map) illustrates that simulated trials consisted of $384$ <span style="color:red"> 18-m </span> $\times$ <span style="color:red"> 73-m </span> "plots," each of which was assigned an N fertilizer application rate. Each plot was made up of four (4-rows $\times$ 1-column) "subplots," which were unit of analysis used in subsequent statistical analysis. Each subplot contained of a 6-rows $\times$ 6-columns grid of thirty-six <span style="color:red"> 3-m </span> $\times$ <span style="color:red"> 3-m </span> "cells." All field characteristics and yield data were generated at the cell level.

```{r field-map, fig.width=4, fig.height=3, fig.cap = "Division of field: plots, subplots and cells"}
# knitr::include_graphics("./GitControlled/Writing/Re_JournalSubmission/field_devision_v2.png")
readPNG("./GitControlled/Writing/Re_JournalSubmission/field_devision_v2.png") %>% grid.raster()
```

@@

```{r variogram-parameters, tab.cap = "The parameters for the variogram modeling"}
variogram_tb
```

@@

```{r fig-parameters, fig.width=6, fig.height=5, fig.cap = "An illustrative example of spatial distributions of field characteristics"}
# field_alpha + field_beta + field_ymax
grid.arrange(field_ymax, field_alpha, field_beta, field_m_error, ncol=2, nrow=2)
```

@@

```{r fig-optN, fig.width=3, fig.height=4, fig.cap = "An illustrative example of a spatial distribution of true optimal N rate"}
vis_optN_cell
```

@@

```{r fig-nExp, fig.width=3, fig.height=4, fig.cap = "An illustrative example of experimental design of N"}
field_Ndesign
```

@@

```{r field-yield-cell, fig.width=3, fig.height=4, fig.cap = "An illustrative example of a yield distribution map at the cell resolution"}
vis_yield_cell
```

@@

```{r field-yield-subplot, fig.width=3, fig.height=4, fig.cap = "An illustrative example of a yield distribution map at the subplot resolution"}
vis_yield_subplot
```

@@

In the third scenario, denoted "aabbyy," the researcher's dataset did not contain values of $\alpha^{i}$, $\beta^{i}$, or $ymax^{i}$. Rather, it contained subplot-level values of permutations of $\alpha$, $\beta$, and $ymax$. Specifically, for each cell $(i,j)$, we assigned one of a set of spatially autocorrelated weights $r_{\alpha}^{i,j} \in (0,1)$ to the $\alpha$ variable, another set of spatially autocorrelated weights $r_{\beta}^{i,j} \in (0,1)$ to the $\beta$ variable, and another set of spatially autocorrelated weights $r_{ymax}^{i,j} \in (0,1)$ to the ${ymax}^{i,j}$ variable. (There was no spatial correlation between the $r_{\alpha}^{i,j}$ and $r_{\beta}^{i,j}$ data, between $r_{\alpha}^{i,j}$ and $r_{ymax}^{i,j}$ data, or between $r_{\beta}^{i,j}$ and $r_{ymax}^{i,j}$.) Six new covariates were created with the weights and the original covariates $\alpha^{i,j}$, $\beta^{i,j}$ and $ymax^{i,j}$. 
These were 
$\alpha_{1}^{i,j} = r_{\alpha}^{i,j} \alpha^{i,j}$, $\alpha_{2}^{i,j} = (1- r_{\alpha}^{i,j}) \alpha^{i,j}$,
$\beta_{1}^{i,j} = r_{\beta}^{i,j} \beta^{i,j}$, $\beta_{2}^{i,j} = (1- r_{\beta}^{i,j}) \beta^{i,j}$,
$ymax_{1}^{i,j} = r_{ymax}^{i,j} ymax^{i,j}$, and $ymax_{2}^{i,j} = (1- r_{ymax}^{i,j}) ymax^{i,j}$.
For each $\alpha$, $\beta$ and $ymax$, these procedures created two new covariates that were highly spatially autocorrelated with the original covariate, which was meant to reflect that many of the soil/field characteristics used to analyze OFPE data are spatially correlated. The values of $\alpha^{i,j}$, $\beta^{i,j}$ and $ymax^{i,j}$ did not indicate specific soil or field characteristics, but were functions of such. In reality, scientists cannot directly observe variables that can accurately predict yield plateau level, soil N content and N uptake efficiency. Rather, they use multiple observed soil/field characteristics to explain such phenological phenomena. For example, instead of using $ymax^{i,j}$ directly, some topographical land features and soil properties such as elevation and soil sand content could be used as yield-limiting factors [@jiang2004effect; @kravchenko2000correlation]. In reality, scientists may include more than three variables as covariates. This scenario should reduce the accuracy of EONR modeling for all the models compared to the ideal case.

@@

## Training Models and and Estimating Site-specific EONRs

@@

The accuracy of EONR estimation was judged based RMSE ($kg/ha$) of EONR estimation against the true EONR at the subplot level.
In addition, using the EONR estimates, we also calculated profit-deficit ($\$/ha$) (i.e., $\hat{\pi}_{def}$) relative to the true maximum profit at the subplot level. The maximized profit is the profit under the true yield response functions evaluated at $N_{opt}^{i}$. For RF, BRF, and CNN, the accuracy of yield prediction was judged based on the RMSE ($kg/ha$) of the yield prediction against the true yields at the subplot level.

@@

```{r table-optN-subplot-wide, tab.cap = "Mean RMSE of EONR (kg/ha) estimation and profit-deficit ($/ha) by ML methods and modeling scenarios"}
report_table_optN
# knit_print(report_table_optN)
```

@@

Interestingly, CNN completely failed in its estimations of site-specific EONRs. This happened because the site-specific yield response functions identified by CNN were linear, and all with the same slope. Therefore, CNN failed to capture a declining marginal product of N. The constant marginal product caused the estimated EONR to take on either the lowest or highest experimental N application rate. To examine whether this problem is a product solely of the specific choice of CNN architecture, we also tested another CNN model, which was analogous to CNN-ST proposed by @barbosa2020modeling. 
This architecture also estimated that yield and N rates were perfectly linearly related, and also estimated site-specific EONR poorly. Note that @barbosa2020modeling just compared models through the lens of yield prediction but did not use the estimated models to derive EONRs. In some cases, CNN predicted yield well; but in all cases it predicting EONR poorly.

@@

```{r dist-optN, fig.width=6, fig.height=7, fig.cap = "Distributions of RMSE of EONR estimation over simulations"}
plot_dis_optN
```

```{r dist-piLoss, fig.width=6, fig.height=7, fig.cap = "Distributions of profit deficit over simulations"}
piLoss_density
```

@@

Figure \@ref(fig:plot-tre) illustrates the underlying cause of CF-base's superior EONR predictions. It plots the estimated treatment effects against true treatment effects by treatment types and ML methods (RF, BRF, and CF-base) under the "aabbyytt" scenario in one of the $1000$ simulations rounds. The figure shows that the estimated treatment effects did not exhibit significant bias, with the points clustered around the red 1-to-1 line. However, they differ substantially in the accuracy of their treatment effect estimations. Points are clustered much more tightly about the 1-to-1 line for CF-base than RF and BRF. That is, CF-base estimates treatment effects much more efficiently than do the RF and BRF methods, which focus on yield prediction rather than treatment effect estimation.

```{r plot-tre, fig.width=7, fig.height=5, fig.cap = "True treatment effects vs estimated treatment effects (scenario: aabbyytt)"}
figure_te
```

@@

```{r plot-y-optN, fig.width=7, fig.height=5, fig.cap = "The relationship in the performance between EONR estimation and yield prediction by ML methods and modeling scenarios"}
fig_y_optN
```

@@

```{r table-y-subplot-wide, tab.cap = "Mean RMSE of yield prediction (kg/ha) by ML methods and modeling scenarios"}
report_table_y
```

@@

```{r count-table, tab.cap = "Relationship between the EONR prediction performances (as measured by profit) and the yield prediction performances for prediction-oriented ML methods"}
report_summary_res_CNN_RF_BRF
```

>>>>

```{r echo = F, results='hide', message=FALSE}
library(knitr)
library(here)

# === set wd ===#
# Notes: not really necessary to set the working directory as long as this Rmd file is part of an R project and the evaluation directory was set to "Project".

opts_knit$set(root.dir = here())
# setwd(here())
# opts_knit$set(root.dir = "~/OneDrive - University of Nebraska-Lincoln/ML_VRA")

@@

```{r, cache = TRUE}
# === packages ===#
library(sf)
library(png)
library(data.table)
library(grf)
library(RColorBrewer)
library(patchwork)
library(plotly)
library(magrittr)
library(ggplot2)
library(viridis)
library(tmap)
library(tidyverse)
library(ggthemes)
library(mgcv)
library(ggpubr)
library(flextable)
library(officer)
library(officedown)
library(modelsummary)
library(raster)
library(bookdown)
```

```{r analyses, results = "hide", fig.show='hide', cache = FALSE}
#/*=================================================*/
#' # Preparation
#/*=================================================*/
# This is where you prepare data sets and results

#/*----------------------------------*/
#' ## Create figures and tablesb
#/*----------------------------------*/
# --- Convert an Rmd file into a pure R file using purl() ---#
knitr::purl(here("GitControlled/Codes/PrepareResults.Rmd"), output = here("GitControlled/Codes/PrepareResults.R"))

# --- Read the simulation results and prepare figures and tables ---#
source(here("GitControlled/Codes/PrepareResults.R"))
```

@@

<!-- + background on nitrogen management (economics, environment, precision agriculture)   -->

<!-- The modern agricultural production system which is featured by the enlargement of fields due to capital-labor substitution, has been confronting the difficulty of considering the spatial and temporal within-field heterogeneity of crop management[@zhang2002precision]. Precision agriculture(PA) is the combination of technologies and practices developed to assist decision-making for such complicated management utilizing the power of information technology[@cook1998precision; @finger2019precision]. One of the leading PA technologies is variable rate application(VRA), and it has been actively investigated for site-specific input management, especially for nitrogen and seed management [@Khosla2002; @shanahan2004feasibility; @ruffo2006site; @koch2004economic; @ping2008site; @shanahan2008responsive; @munnaf2020site ;@munnaf2021map].
Efficient management of N fertilizer is a critical component of crop production from the agronomic, economic, and environmental standpoints. Inappropriate amount or timing of N application adversely affects both biomass yields and grain quality for some crops. Furthermore, inefficient use of N fertilizer not only results in economic loss but causes ambient environmental degradation[@blumenthal2008importance; @dinnes2002nitrogen; @tremblay2006adapting]. \ -->

@@

CF was developed specifically for identifying heterogeneous treatment effects [@athey2016recursive; @Wager2018a]. Like RF, CF works by building an ensemble of trees, called *causal trees* (CT), by recursively partitioning covariate space into two leaves, where each leaf estimates conditional average treatment effects as the mean differences between the dependent variable (here, yield) of the treated and control observations that belong to the leaf. Figure \@ref(fig:ct-ex) shows an example of CT where the treatment is an increase in seed rate from $27,000$ seeds/ha (control) to $31,000$ seeds/ha (treated) with apparent electrical conductivity ($\textit{ecs}$) and topographical $\textit{slope}$ as variables to explain the heterogeneity of the impact of the treatment impact. 
In the first node, all the samples ($1024$ observations) belong to the node, and the causal effect of the treatment was calculated as the difference in the mean yield of the treated and control groups, which turned out to be $593$ kg/ha in this example.

```{r ct-ex, fig.width=4, fig.height=4, fig.align="center", fig.cap = "An Illustrative Example of Causal Tree"}
knitr::include_graphics(here("GitControlled/Writing/JournalSubmission/ctree_ex.png"))
```

@@

Monte Carlo simulations were conducted for nitrogen experiments for corn production to estimate the economically optimal nitrogen rate (EONR). The first step in our simulation was to generate one thousand synthetic fields that "resembled" an actual field. For each simulated field, we generated a training data set and a validation data set. Field characteristics and yield were simulated based on the Mitscherlich-Baule (MB) yield response function. We used the RF, BRF, and CNN ML methods to estimate yield response functions, and we used the CF method to estimate treatment effects. For each method, we estimated site-specific EONRs using the training data set, and checked the statistical performance of the predicted site-specific EONR using the validation data set. Detailed descriptions of these steps are provided below. The R codes that implement the simulation analyses are available at https://github.com/tmieno2/ML_VRA.git.

@@

### Simulated Experimental Fields
<!-- experimental field: area: 51ha, plot: 384, subplot:1536, cell:55296   -->
<!-- field with padding: area: 57ha, plot: 476, subplot:1700, cell:61200   -->

The simulated field was created based on an actual 51-ha Illinois field on which the Data-Intensive Farm Management (DIFM) project [@bullock2019] has run multiple OFPEs. Figure \@ref(fig:field-map) illustrates that simulated trials consisted of $384$ 60-ft $\times$ 240-ft "plots," each of which was assigned an N fertilizer application rate. Each plot was made up of four (4-rows $\times$ 1-column) "subplots," which were unit of analysis used in subsequent statistical analysis. Each subplot contained of a 6-rows $\times$ 6-columns grid of thirty-six 10-ft $\times$ 10-ft "cells." All field characteristics and yield data were generated at the cell level.

```{r field-map, fig.width=4, fig.height=3, fig.cap = "Division of Field: plots, subplots and cells"}
knitr::include_graphics(here("GitControlled/Writing/JournalSubmission/field_devision.png"))

```

@@

```{r variogram-parameters, tab.cap = "The Parameters for the Variogram Modeling"}
variogram_tb
```

@@

```{r fig-parameters, fig.width=6, fig.height=5, fig.cap = "An Illustrative Example of Spatial Distributions of Field Characteristics"}
# field_alpha + field_beta + field_ymax
grid.arrange(field_ymax, field_alpha, field_beta, field_m_error, ncol=2, nrow=2)
```

@@

```{r fig-optN, fig.width=3, fig.height=4, fig.cap = "An Illustrative Example of a Spatial Distribution of True Optimal N Rate"}
vis_optN_cell
```

@@

```{r fig-nExp, fig.width=3, fig.height=4, fig.cap = "An Illustrative Example of Experimental Design of N"}
field_Ndesign
```

@@

```{r field-yield-cell, fig.width=3, fig.height=4, fig.cap = "An Illustrative Example of a Yield Distribution Map at the Cell resolution"}
vis_yield_cell
```

@@

```{r field-yield-subplot, fig.width=3, fig.height=4, fig.cap = "An Illustrative Example of a Yield Distribution Map at the Subplot resolution"}
vis_yield_subplot
```

@@

In the third scenario, denoted "aabbyy," the researcher's dataset did not contain values of $\alpha^{i}$, $\beta^{i}$, or $ymax^{i}$. Rather, it contained subplot-level values of permutations of $\alpha$, $\beta$, and $ymax$. Specifically, for each cell $(i,j)$, we assigned one of a set of spatially autocorrelated weights $r_{\alpha}^{i,j} \in (0,1)$ to the $\alpha$ variable, another set of spatially autocorrelated weights $r_{\beta}^{i,j} \in (0,1)$ to the $\beta$ variable, and another set of spatially autocorrelated weights $r_{ymax}^{i,j} \in (0,1)$ to the ${ymax}^{i,j}$ variable. (There was no spatial correlation between the $r_{\alpha}^{i,j}$ and $r_{\beta}^{i,j}$ data, between $r_{\alpha}^{i,j}$ and $r_{ymax}^{i,j}$ data, or between $r_{\beta}^{i,j}$ and $r_{ymax}^{i,j}$.) Six new covariates were created with the weights and the original covariates $\alpha^{i,j}$, $\beta^{i,j}$ and $ymax^{i,j}$. 
These were 
$\alpha_{1}^{i,j} = r_{\alpha}^{i,j} \alpha^{i,j}$, $\alpha_{2}^{i,j} = (1- r_{\alpha}^{i,j}) \alpha^{i,j}$,
$\beta_{1}^{i,j} = r_{\beta}^{i,j} \beta^{i,j}$, $\beta_{2}^{i,j} = (1- r_{\beta}^{i,j}) \beta^{i,j}$,
$ymax_{1}^{i,j} = r_{ymax}^{i,j} ymax^{i,j}$, and $ymax_{2}^{i,j} = (1- r_{ymax}^{i,j}) ymax^{i,j}$.
For each $\alpha$, $\beta$ and $ymax$, these procedures created two new covariates that were highly spatially autocorrelated with the original covariate, which was meant to reflect that many of the soil/field characteristics used to analyze OFPE data are spatially correlated. The values of $\alpha^{i,j}$, $\beta^{i,j}$ and $ymax^{i,j}$ did not indicate specific soil or field characteristics, but were functions of such. In reality, scientists cannot directly observe variables that can accurately predict yield plateau level, soil N content and N uptake efficiency. Rather, they use multiple observed soil/field characteristics to explain such phenological phenomena. For example, instead of using $ymax^{i,j}$ directly, some some topographical land features and soil properties such as elevation and soil sand content could be used as yield-limiting factors [@jiang2004effect; @kravchenko2000correlation]. In reality, scientists may include more than three variables as covariates. This scenario should reduce the accuracy of EONR modeling for all the models compared to the ideal case.

@@

## OFPE Experiments and Estimating Site-specific Optimal N Rates

@@

The accuracy of EONR estimation was judged based RMSE ($kg/ha$)of EONR estimation against the true EONR at the subplot level.
In addition, using the EONR estimates, we also calculated profit-deficit ($\$/ha$) (i.e., $\hat{\pi}_{def}$) relative to the true maximum profit at the subplot level. The maximized profit is the profit under the true yield response functions evaluated at $N_{opt}^{i}$. For RF, BRF, and CNN, the accuracy of yield prediction was judged based on the RMSE ($kg/ha$) of the yield prediction against the true yields at the subplot level.

@@

```{r table-optN-subplot-wide, tab.cap = "Mean RMSE of EONR Estimation and Profit-deficit by ML Methods and Modeling Scenarios"}
report_table_optN
# knit_print(report_table_optN)
```

@@

Interestingly, CNN completely failed in its estimations of site-specific EONRs. This happened because the site-specific yield response functions identified by CNN were linear, and all with the same slope. Therefore, CNN failed to capture a declining marginal product of N. The constant marginal product caused the estimated EONR to take on either the lowest or highest experimental N application rate. To examine whether this problem is product solely of the specific choice of CNN architecture, we also tested another CNN model, which was analogous to CNN-ST proposed by @barbosa2020modeling. 
This architecture also estimated that yield and N rates were perfectly linearly related, and also estimated site-specific EONR poorly. Note that @barbosa2020modeling just compared models through the lens of yield prediction but did not use the estimated models to derive EONRs. In some cases, CNN predicted yield well; but in all cases it predicting EONR poorly.

@@

```{r dist-optN, fig.width=6, fig.height=7, fig.cap = "Distributions of RMSE of EONR Estimation over Simulations"}
plot_dis_optN
```

```{r dist-piLoss, fig.width=6, fig.height=7, fig.cap = "Distributions of Profit Deficit over Simulations"}
piLoss_density
```

@@

Figure \@ref(fig:plot-tre) illustrates the underlying cause of CF-base's superior EONR predictions. It plots the estimated treatment effects against true treatment effects by treatment types and ML methods (RF, BRF, and CF-base) under the "aabbyytt" scenario in one of the $1000$ simulations rounds. The figure shows that the estimated treatment effects did not exhibit significant bias, with the points clustered around the red 1-to-1 line. However, they differ substantially in the accuracy of their treatment effect estimations. Points are clustered much more tightly about the 1-to-1 line much for CF-base than RF and BRF. That is, CF-base estimates treatment effects much more efficiently than do the RF and BRF methods, which focus on yield prediction rather than treatment effect estimation.

```{r plot-tre, fig.width=7, fig.height=5, fig.cap = "True Treatment Effects vs Estimated Treatment Effects (Scenario: aabbyytt)"}
figure_te
```

@@

```{r plot-y-optN, fig.width=7, fig.height=5, fig.cap = "The Relationship in the Performance between EONR Estimation and Yield Prediction by ML Methods and Modeling Scenarios"}
fig_y_optN
```

@@

```{r table-y-subplot-wide, tab.cap = "Mean RMSE of Yield Prediction by ML Methods and Modeling Scenarios"}
report_table_y
```

@@

```{r count-table, tab.cap = "Relationship between the EONR Prediction Performances (as measured by profit) and the Yield Prediction Performances for Prediction-oriented ML Methods"}
report_summary_res_CNN_RF_BRF
```
